%TC:ignore

\begin{abstract}
This interim report details the 4-month progress on a project to design, implement, and verify, a multi-core FPGA RISC processor. The project has been split into two stages: firstly to build a functional single-core RISC processor, and then secondly to add multiprocessor principles and functionality to it.

Current multiprocessor and network-on-chip communication methods have been discussed and how they could be included in this multi-core RISC design. To-date, a 16-bit instruction set architecture has been designed featuring common load/store instructions, comparison, and bitwise operations. A single-core processor has been implemented in Verilog and verified using simulations/test benches running various simple software programs.

Future tasks have been planned and will focus on the second stage of the project. Work will start on designing a loosely coupled multiprocessor communication interface and bringing them to the single-core processor.
\end{abstract}

\chapter*{Declaration of Academic Integrity}
%\addcontentsline{toc}{chapter}{Declaration of Academic Integrity}
The candidate confirms that the work submitted is his/her own, except where work which has formed part of jointly-authored publications has been included. The contribution of the candidate and the other authors to this work has been explicitly indicated in the report. The candidate confirms that appropriate credit has been given within the report where reference has been made to the work of others.

This copy has been supplied on the understanding that no quotation from the report may be published without proper acknowledgement. The candidate, however, confirms his/her consent to the University of Leeds copying and distributing all or part of this work in any forms and using third parties, who might be outside the University, to monitor breaches of regulations, to verify whether this work contains plagiarised material, and for quality assurance purposes.

The candidate confirms that the details of any mitigating circumstances have been submitted to the Student Support Office at the School of Electronic and Electrical Engineering, at the University of Leeds.
\vfill

\noindent 
Name: Ben David Lancaster \\
Date: \today
\newpage

\chapter*{Acknowledgements}
I would like to thank my supervisor, David Cowell, and assessor, David Moore, for giving me the opportunity to explore a project of my own choosing, and for their support and guidance throughout this project.


\if 0
\clearpage
\section*{Revision History}
\begin{table}[h]
\def\arraystretch{1.3}
\begin{tabularx}{\textwidth}{|l|l|X|}
\hline
Date & Version & Changes \\
%\arrayrulecolor{\color{red}}
\specialrule{2pt}{-2pt}{0pt}
10/04/2019 & 2.02 & Update future stages. \\ \hline
05/04/2019 & 2.01 & Fix processor RTL diagram. \\ \hline
04/04/2019 & 2.00 & Initial processor RTL diagram. \\ \hline
01/04/2019 & 1.00 & Initial section outline. \\ \hline
\end{tabularx}
\caption*{Document revisions.}
\end{table}
\fi


%TC:endignore